---
title: "Domestic citation analysis"
author: "Zuguang Gu ( z.gu@dkfz.de )"
date: '`r Sys.Date()`'
output: 
  html_document:
    toc: true
    fig_caption: true
---

```{r, echo = FALSE, message = FALSE}
library(knitr)
knitr::opts_chunk$set(
    error = FALSE,
    tidy  = FALSE,
    message = FALSE,
    warning = FALSE,
    fig.width = 6, fig.height = 6,
    fig.align = "center"
)
```

We first load the results of domestic citation analysis.

```{r}
df2 = readRDS("domestic_enrichment.rds")
head(df2)
```

We will use the following columns:

- `citations`: The number of citations that a country A cites itself, denoted as $k_{A-A}$.
- `total_cited`: The number of total citations that A is cited globally, denoted as $m_{A-}$.
- `total_citing`: The number of total citations that A cites globally, denoted as $n_{-A}$.
- `global_citation`: The total number of citations in the globe, denoted as $N$.
- `expected`: The expected number of citations that A cites itself, calculated as $m_{A-}n_{-A}/N$.
- `log2_fc`: The enrichment score.


Emperically, $k_{A-A}$ has linear relations to $n_{-A}$ and $m_{A-}$, and it can be demonstrated by the following
two scatter plots (each data point corresponds to one country):

$$
\begin{align*}
\log_2 n_{-A} &= a_1 \cdot \log_2 k_{A-A} + b_1 \\
\log_2 m_{A-} &= a_2 \cdot \log_2 k_{A-A} + b_2 \\
\end{align*}
$$


```{r, fig.width = 10}
par(mfrow = c(1, 2))
plot(df2$citations, df2$total_citing, log = "xy", xlab = "# Domestic citations", ylab = "# Citations A cites")
plot(df2$citations, df2$total_cited,  log = "xy", xlab = "# Domestic citations", ylab = "# Citations A is cited")
```

We can fit linear regression models on the log2-scale:

```{r}
summary(lm(log2(df2$total_citing) ~ log2(df2$citations)))
summary(lm(log2(df2$total_cited) ~ log2(df2$citations)))
```

In the manuscript, we demonstrated there is a negative linear relation between the enrichment scores and the number
of domestic citations:

$$
\begin{align*}
s_{A-A} &= \log_2 k_{A-A} - \log_2 n_{-A} + \log_2 N - \log_2 m_{A-} \\
  &= (1 - a_1 - a_2) \log_2 k + \log_2 N - b_1 - b_2 \\
  &= -0.70 \log_2 k_{A-A} + 16.41
\end{align*}
$$


```{r}
plot(df2$citations, df2$log2_fc, log = "x", xlab = "# Domestic citations", ylab = "Enrichment scores")
```

In this document, we will prove when there are positive enrichment, $a_1 + a_2 > 1$ and it results in the decrease
of the enrichment scores.

First let's simplify the notations: $k_{A-A}$ to $k$, $m_{A-}$ to $m$ and $n_{-A}$ to $n$. From the two linear fits, we have:


$$
\begin{align*}
n &= k^{a_1} \cdot 2^{b_1} \\
m &= k^{a_2} \cdot 2^{b_2} \\
mn &= k^{a_1 + a_2} \cdot 2^{b_1 + b_2} \\
\end{align*}
$$

Denote $k_\mathrm{exp}$ as the expected number of citations, and $mn$ can also be calculated as:  

$$
mn = k_\mathrm{exp}N
$$

It is easy to see, when there is no enrichment of domestic citations, $a_1 + a_2 = 1$.

Now we have:

$$
k^{a_1 + a_2} \cdot 2^{b_1 + b_2} = k_\mathrm{exp}N
$$

Since all the countries share the same coeffcients from the linear fit, we can simply treat different countries
correspond to different stages of the citation network expansion, where small countries correspond to early states.
Then for two countries with $k_2 > k_1$ and $k_{\mathrm{exp},2} > k_{\mathrm{exp},1}$, we divide the previous equation on both sides:

$$
\begin{align*}
\left( \frac{k_2}{k_1} \right)^{a_1 + a_2} &= \frac{k_{\mathrm{exp},2}}{k_{\mathrm{exp},1}} \\
(a_1 + a_2)\cdot \log(k_2/k_1) &= \log(k_{\mathrm{exp},2}/k_{\mathrm{exp},1}) \\
a_1 + a_2 &= \frac{\log(k_{\mathrm{exp},2}/k_{\mathrm{exp},1})}{\log(k_2/k_1)} \\
\end{align*}
$$

Denote $T_{A-A}$, $T_{A-}$ and $T_{-A}$ as the citation networks constructed by the citation relations in the three groups (A-domestic, A-citing
and A-cited). Let's assume from country 1 to 2, $T_{A-}$ and $T_{-A}$ both expands by a factor of $\alpha$ ($\alpha > 1$), i.e. $m_2 = \alpha \cdot m_1$ and $n_2 = \alpha \cdot n_1$. Then the expansion rate of the citation network if there is no enrichment is

$$
\frac{k_{\mathrm{exp},2}}{k_{\mathrm{exp},1}} = \frac{m_2n_2/N}{m_1n_2/N} = \alpha^2
$$

Next we look at the expansion rate of $k_2/k_1$. If there is an over-representation, $k$ can be decomposed to

$k = k_\mathrm{exp} + k_\mathrm{diff}$

where $k_\mathrm{diff}$ is the over-represented portion of $k$. It can be expected that when $m$ or $n$ increases, $k_\mathrm{diff}$
also increases. Now the question is does it increase by a fixed a rate or does it still depends on $m$ or $n$ (i.e. increase faster when $m$ or $n$ gets larger?)

We next check how $k_\mathrm{diff}/m$ or $k_\mathrm{diff}/n$ changes to the size of $T_{A-}$ and $T_{-A}$, i.e. 


```{r, fig.width = 10}
diff = df2$citations - df2$expected
par(mfrow = c(1, 2))
plot(df2$total_cited, diff/df2$total_cited, log = "x", xlab = "# Citations A is cited (m)", ylab = "diff/m")
plot(df2$total_citing, diff/df2$total_citing, log = "x", xlab = "# Citations A cites (n)", ylab = "diff/n")
```

We can see, in both plots, there is a positive linear relations in the log-scale. Let's fit the two linear models.
Note China (index = 11) and United States (index = 80) are removed from the fitting because they are outliers on the scatter plots.

```{r}
p1 = diff/df2$total_citing
summary(lm(p1[-c(11, 80)] ~ log(df2$total_citing[-c(11, 80)])))

p2 = diff/df2$total_cited
summary(lm(p2[-c(11, 80)] ~ log(df2$total_cited[-c(11, 80)])))
```



Both slope and intercept coeffcients are very small. To reduce the complexity of the followig equations, we assume $m = n$. 
We discuss the following two scenarios.

**Scenario 1**: there is a weak positive relations between $p$ and $m$ on the log-scale, thus on the normal scale, we can approximate
$p$ as a constant and not dependent to $m$, i.e.:

$$
\begin{align*}
k_\mathrm{diff} &= pm \\
k &= k_\mathrm{exp} + pm \\
\end{align*}
$$

Then we can calculate $k_2/k_1$ as:

$$
\begin{align*}
\frac{k_2}{k_1} &= \frac{k_{\mathrm{exp},2} + p m_2}{k_{\mathrm{exp},1} + p m_1} \\
 &= \frac{\alpha^2 k_{\mathrm{exp},1} + p \alpha m_1}{k_{\mathrm{exp},1} + p m_1} \\
  &= \frac{\alpha^2 (k_{\mathrm{exp},1} + p m_1 ) + p \alpha m_1 - p \alpha^2 m_1}{k_{\mathrm{exp},1} + p m_1} \\
  &= \alpha^2 - \frac{p \alpha(\alpha - 1) m_1}{k_{\mathrm{exp},1} + p m_1} \\ 
\end{align*}
$$

Note since $\alpha > 1$, we have $k_2/k_1 < \alpha^2$.

In the equation, it also shows the expansion of the citation network can be decomposed into two parts, the part without 
citation enrichment which grows in quadratic speed ($\sim \alpha^2$), and the "over-represented" part which grows in linear
speed ($\sim \alpha$). Mixing these two speeds results the final speed being weak than quadratic.


**Scenario 2**: let's assume there is the following relation where $p$ is dependent on $\log m$. Note for simplicity we 
removed the intercept (the value of the intercept is very small and it can be ignored).

$$
p = w \log(m)
$$

where $w$ is slop coeffcient from the liear fit.

Then $k_2/$k_1$ can be written as:

$$
\begin{align*}
\frac{k_2}{k_1} &= \frac{k_{\mathrm{exp},2} + p_2 \cdot m_2}{k_{\mathrm{exp},1} + p_1 \cdot m_1} \\
 &= \frac{\alpha^2 \cdot k_{\mathrm{exp},1} + (w\log (\alpha m_1)) \cdot \alpha m_1}{k_{\mathrm{exp},1} + (w\log m_1) \cdot m_1} \\
  &= \frac{\alpha^2 (k_{\mathrm{exp},1} + w m_1 \log m_1 ) - \alpha^2 w m_1\log m_1 + \alpha w m_1 \log \alpha  + \alpha w m_1 \log m_1}{k_{\mathrm{exp},1} + w m_1 \log m_1} \\
  &= \alpha^2 - \frac{\alpha w m_1 \left[ (\alpha - 1) \log m_1 - \log \alpha \right]}{k_{\mathrm{exp},1} + w m_1 \log m_1} \\ 
\end{align*}
$$

where $\log m_1$ is much larger than $\log \alpha$, which also results in $k_2/k_1 < \alpha^2$.

In this case, in the expansion of the citation network, the "over-represented" part grows in a speed of $\sim \alpha \log \alpha$, but
still slower than the quadratic.

For both scenarios, we can have:

$$
\begin{align*}
k_{\mathrm{exp},2}/ k_{\mathrm{exp},1} &> k_2 / k_1 > 1 \\
\log(k_{\mathrm{exp},2}/ k_{\mathrm{exp},1}) &> \log( k_2 / k_1 ) > 0 \\
\end{align*}
$$

And eventually 

$$
a_1 + a_2 = \frac{\log(k_{\mathrm{exp},2}/k_{\mathrm{exp},1})}{\log(k_2/k_1)} > 1
$$


<br>
<br>
<br>
<br>
<br>
<br>
<br>
<br>
<br>
