---
title: "Visualize the grouping patterns of countries"
author: "Zuguang Gu ( z.gu@dkfz.de )"
date: '`r Sys.Date()`'
output: 
  html_document:
    toc: true
    fig_caption: true
---

```{r, echo = FALSE, message = FALSE}
library(knitr)
knitr::opts_chunk$set(
    error = FALSE,
    tidy  = FALSE,
    message = FALSE,
    warning = FALSE,
    fig.width = 12, fig.height = 6,
    fig.align = "center"
)
```


In the manuscript, countries are clustered into 7 groups based on their influence patterns by a graph clustering method. In this
document, we will apply dimension reduction methods to validate the clustering results.

The influence matrix contains `NA` values where two countries do not have enough number of citations for calculating the influence value.
Also there might be "outliers" when calculating the distance between two countries. Thus, we first defined a "robust distance function"
which removes `NA` values and also extreme values.

```{r}
robust_dist_pair = function(x, y, trim = 0.1) {
	x_mid = mean(x, na.rm = TRUE)
	y_mid = mean(y, na.rm = TRUE)

	dd = sqrt( (x-x_mid)^2 + (y-y_mid)^2 )

	q = quantile(dd, c(trim/2, 1-trim/2), na.rm = TRUE)
	l = dd >= q[1] & dd <= q[2]
	l[is.na(l)] = FALSE
	sqrt( sum( (x[l] - y[l])^2 ) )/sum(l)
}

# by rows
robust_dist = function(m, trim = 0.1) {
	n = nrow(m)
	d = matrix(0, nrow = n, ncol = n)
	rownames(d) = colnames(d) = rownames(m)

	for(i in 1:(n-1)) {
		for(j in (i+1):n) {
			d[i, j] = d[j, i] = robust_dist_pair(m[i, ], m[j, ], trim = trim)
		}
	}

	as.dist(d)
}
```


We load the data. `m` is the influence matrix of 72 countries. `mem` is the clustering results from the graph clustering analysis.

```{r}
m = readRDS("mat_influence.rds")
mem = readRDS("country_groups.rds")

identical(rownames(m), names(mem))
identical(colnames(m), names(mem))
split(names(mem), mem)
```

We will try the following dimension reduction methods: MDS (multidimensional scaling), UMAP and t-SNE. For each method, the distance
will be calculated on rows and columns of `m` separately, thus there are two plots for each method.

```{r}
library(ggplot2)
library(ggrepel)
library(cowplot)
library(grid)
```

```{r}
d = robust_dist(m)
loc = cmdscale(d)

loc = as.data.frame(loc)[, 1:2]
colnames(loc) = c("V1", "V2")
loc$label = names(mem)
loc$group = paste0("group-", mem)

p1 = ggplot(loc, aes(x = V1, y = V2, color = group, label = label)) + 
	geom_point() + geom_text_repel(max.overlaps = Inf, size = 3, show.legend = FALSE) +
	labs(x = "MDS 1", y = "MDS 2") +
	ggtitle("As cited countries / how they influence other countries")
p1 = grid.grabExpr(print(p1))


# the second plot, note `m` is transposed
d = robust_dist(t(m))
loc = cmdscale(d)

loc = as.data.frame(loc)[, 1:2]
colnames(loc) = c("V1", "V2")
loc$label = names(mem)
loc$group = paste0("group-", mem)

p2 = ggplot(loc, aes(x = V1, y = V2, color = group, label = label)) + 
	geom_point() + geom_text_repel(max.overlaps = Inf, size = 3, show.legend = FALSE) +
	labs(x = "MDS 1", y = "MDS 2") +
	ggtitle("As citing countries / how they are influenced by other countries")
p2 = grid.grabExpr(print(p2))

plot_grid(p1, p2, nrow = 1)
```


Next we apply UMAP and t-SNE dimension reduction methods. Note they need a complete matrix with no `NA` value.
Thus, we first impute the `NA` values with the **impute** package.

```{r}
library(impute)
m2 = impute.knn(m)$data
```

We apply a wrapper function `dimension_reduction()` from the **cola** package for UMAP and t-SNE analysis.
Note `dimension_reduction()` is applied on columns.

```{r, fig.keep = "none"}
library(cola)
set.seed(123)
# on the rows of `m2`
loc = dimension_reduction(t(m2), method = "UMAP")

loc = as.data.frame(loc)[, 1:2]
colnames(loc) = c("V1", "V2")
loc$label = names(mem)
loc$group = paste0("group-", mem)
```

```{r}
p1 = ggplot(loc, aes(x = V1, y = V2, color = group, label = label)) + 
	geom_point() + geom_text_repel(max.overlaps = Inf, size = 3, show.legend = FALSE) +
	labs(x = "UMAP 1", y = "UMAP 2") +
	ggtitle("As cited countries / how they influence other countries")
p1 = grid.grabExpr(print(p1))
```

```{r, fig.keep = "none"}
# on the columns of m2
loc = dimension_reduction(m2, method = "UMAP")
loc = as.data.frame(loc)[, 1:2]
colnames(loc) = c("V1", "V2")
loc$label = names(mem)
loc$group = paste0("group-", mem)
```

```{r}
p2 = ggplot(loc, aes(x = V1, y = V2, color = group, label = label)) + 
	geom_point() + geom_text_repel(max.overlaps = Inf, size = 3, show.legend = FALSE) +
	labs(x = "UMAP 1", y = "UMAP 2") +
	ggtitle("As citing countries / how they are influenced by other countries")
p2 = grid.grabExpr(print(p2))

plot_grid(p1, p2, nrow = 1)
```

And the dimension reduction by t-SNE.

```{r, fig.keep = "none"}
# on the rows of `m2`
loc = dimension_reduction(t(m2), method = "t-SNE")

loc = as.data.frame(loc)[, 1:2]
colnames(loc) = c("V1", "V2")
loc$label = names(mem)
loc$group = paste0("group-", mem)
```

```{r}
p1 = ggplot(loc, aes(x = V1, y = V2, color = group, label = label)) + 
	geom_point() + geom_text_repel(max.overlaps = Inf, size = 3, show.legend = FALSE) +
	labs(x = "t-SNE 1", y = "t-SNE 2") +
	ggtitle("As cited countries / how they influence other countries")
```

```{r, fig.keep = "none"}
# on the columns of `m2`
loc = dimension_reduction(m2, method = "t-SNE")

loc = as.data.frame(loc)[, 1:2]
colnames(loc) = c("V1", "V2")
loc$label = names(mem)
loc$group = paste0("group-", mem)
```

```{r}
p2 = ggplot(loc, aes(x = V1, y = V2, color = group, label = label)) + 
	geom_point() + geom_text_repel(max.overlaps = Inf, size = 3, show.legend = FALSE) +
	labs(x = "t-SNE 1", y = "t-SNE 2") +
	ggtitle("As citing countries / how they are influenced by other countries")
p2 = grid.grabExpr(print(p2))

plot_grid(p1, p2, nrow = 1)
```
