---
title: "Mutual influence between West-1 to other groups of countries"
author: "Zuguang Gu ( z.gu@dkfz.de )"
date: '`r Sys.Date()`'
output: 
  html_document:
    toc: true
    fig_caption: true
---

```{r, echo = FALSE, message = FALSE}
library(knitr)
knitr::opts_chunk$set(
    error = FALSE,
    tidy  = FALSE,
    message = FALSE,
    warning = FALSE,
    fig.align = "center"
)
```


We load the data. `m` is the influence matrix between 72 countries. `mem` is the clustering results from the graph clustering analysis.

```{r}
library(ComplexHeatmap)
library(circlize)

m = readRDS("mat_influence.rds")
mem = readRDS("country_groups.rds")

lt = split(names(mem), mem)
lt
```

To be the same as in the manuscript, we extract countries in West-1 (group-1), and perform 
the Louvain graph clustering method to partition West-1 into three subgroups:

```{r}
cn = lt[["2"]]
subm = m[cn, cn]
subm2 = subm
subm2[is.na(subm2)] = 0
subm2[subm2 < 0] = 0
library(igraph)
g2 = graph_from_adjacency_matrix(subm2, mode = "plus", weighted = TRUE)

set.seed(666)
mem2 = membership(cluster_louvain(g2, weight = sqrt(E(g2)$weight), resolution = 1.1))
mem2 = factor(mem2, levels = c(3, 1, 2))
levels(mem2) = c("Northern Europe", "English-speaking", "Mainland Europe")
mem2
```

And the heatmap the same as Figure 7 in the main manuscript:

```{r, fig.width = 6, fig.height = 5.5}
col_fun = colorRamp2(c(-2, 0, 2), c("blue", "white", "red"))
ht = Heatmap(subm, name = "Enrichment score",
	row_title = "Cited/influencing countries",
	column_title = "Citing/influenced countries", column_title_side = "bottom",
	col = col_fun,
	row_split = mem2, column_split = mem2,
	na_col = "#DDDDDD",
	cluster_row_slices = FALSE, cluster_column_slices = FALSE,
	show_row_dend = FALSE, show_column_dend = FALSE,
	border = "black",
	width = unit(4, "mm")*nrow(subm),
	height = unit(4, "mm")*ncol(subm)
)

draw(ht, column_title = "Influence map only in West-1 countries",
	heatmap_legend_list = list(Legend(at = "Not available", legend_gp = gpar(fill = "#DDDDDD"))))
```

We remove Estonia and Luxembourg because most of the enrichment values are missing between they two and other
countries in World-2.

```{r}
cn = setdiff(cn, c("Estonia", "Luxembourg"))
mem2 = mem2[!names(mem2) %in% c("Estonia", "Luxembourg")]
mem2_col = c("Northern Europe" = 2, "English-speaking" = 3, "Mainland Europe" = 4)
```

The first heatmap shows how World-1 countries influence World-2.

```{r, fig.width = 10, fig.height = 5}
cn2 = setdiff(names(mem), c(lt[["2"]], "United States"))
Heatmap(m[cn, cn2], name = "Enrichment score",
	col = col_fun,
	right_annotation = rowAnnotation(group = mem2, col = list(group = mem2_col)),
	column_title = "How World-1 influences World-2",
	column_split = mem[cn2]
)
```

The second heatmap shows how World-1 is influenced by World-2.

```{r, fig.width = 6, fig.height = 10}
draw(Heatmap(m[cn2, cn], name = "Enrichment score",
	col = col_fun,
	top_annotation = HeatmapAnnotation(group = mem2, col = list(group = mem2_col)),
	column_title = "How World-1 is influenced by World-2",
	row_split = mem[cn2]
), merge_legends = TRUE)
```

The classification of "Northern Europe/English-speaking/Mainland Europe" cannot be observed in both heatmaps.

